% !TEX Program = xelatex
\documentclass[print, master, vlined]{DissertUESTC}
% \documentclass[print, master, review, vlined]{DissertUESTC}
% Keep bibliography fully justified while improving line breaks for URLs and English titles.
\usepackage{xurl}
\makeatletter
\renewcommand{\bibfont}{%
	\zihao{5}\setlength{\baselineskip}{20bp}\selectfont%
	\tolerance=2000%
	\emergencystretch=2em%
	\hyphenpenalty=200%
	\exhyphenpenalty=100%
}
\g@addto@macro\UrlBreaks{\do\/\do-\do\_\do\.\do\?\do\&\do\=\do\#}
\Urlmuskip=0mu plus 1mu minus 1mu\relax
\makeatother

\begin{document}
	
	% 以下两条命令为高亮示例文档中的关键内容而设，正式撰写时切不可使用
	\newcommand{\shad}[1]{\textcolor{DodgerBlue}{\ttfamily #1}}
	\newcommand{\shadcmd}[1]{\shad{$\backslash$#1}}

	% \raggedbottom  % 此命令可让LaTeX像Word那样直接堆叠页面内容，而不再默认拉伸段间距，代价是页尾可能会有明显空白
	

	% 封面示例————非双学位学士论文
	%	\uestccover[<学院名称排版风格>]{<论文题目>}{<学科专业>}{<学号>}{<作者姓名>}{<指导教师>}{<教师职称>}{<学院>} % 单学位学士、研究生通用
	\uestccover{面向大规模低轨卫星的在网计算研究}{信息与通信工程}{202321011337}{龚科成}{孙罡}{教\hspace{1em}授}{信息与通信工程学院}  % 空封面
				
	% 中文扉页
	\ClsNum{}
	\ClsLv{公开}
	\UDC{}
	\DissertationTitle{面向大规模低轨卫星的在网计算研究}
	\Author{龚科成}
	\Supervisor{孙罡}{教授}{电子科技大学}{成都}
	\Major{信息与通信工程}
	\Date{}{}
	\Grant{电子科技大学}{}
	\Reviewer{}{}
	\uestczhtitlepage

	% 英文扉页
	\uestcentitlepage{Research on In-Network Computing for Large-Scale Low Earth Orbit Satellites}
					{Information and Communication Engineering}
					{202321011337}
					{Gong Kecheng}
					{Prof. Sun Gang}
					{}
					{School of Information and Communication Engineering}

	% 独创性声明与论文使用授权页
	\declaration{}{}{}

	
	
	% 开启中文摘要
	\zhabstract
	
	大规模低轨（LEO）卫星巨型星座凭借低时延、广覆盖和全球互联能力，正逐步成为天地一体化信息网络的重要基础设施。然而，与地面骨干网相比，LEO 星座的星间链路带宽更有限，且卫星轨道、拓扑结构和部分链路信息具有较强的公开性与可预测性，使攻击者能够更容易构造面向关键链路的链路洪泛攻击（LFA）。这类攻击通过操纵大量分布式节点向合法目标发送低速但协同的流量，使攻击流在目标链路上汇聚形成拥塞，从而显著降低合法业务的吞吐并破坏网络可用性。

	现有防御方法仍难以直接适用于 LEO 场景。一方面，依赖地面控制器或地面清洗中心的集中式方案受制于星地控制环路时延，难以及时响应毫秒级动态变化的攻击；另一方面，面向地面网络设计的可编程数据面方案通常假设拓扑稳定、硬件资源充足，并多依赖聚合速率等单一特征，在多 decoy 分摊、低速化和脉冲化等隐蔽攻击下容易出现漏检。此外，仅在受害链路附近进行本地缓解，难以阻止攻击流在上游链路上的无效传输。针对上述问题，本文面向星载可编程数据面受限环境，研究了一种从单节点检测到多节点协同推回的在网计算防御方法。

	针对单星场景下隐蔽 LFA 难以准确识别的问题，本文提出多签名在网检测与队列缓解方法 MS-SatShield。该方法采用“Top-$k$ 候选生成+多维结构特征提取”的级联式检测架构，在有限状态开销下联合利用聚合速率、扇入度（FI）、扇出度（FO）和持续性（Persist）等特征，实现对退化攻击的鲁棒检测。实验结果表明，在多 decoy 分摊退化设定下（$M=100$），仅依赖聚合速率的基线方案 F1 降至 0.017，而引入 FI/FO 后 F1 恢复至 0.908；在脉冲式 on-off 攻击下，持续性特征进一步将 Recall 提升至 0.999，且数据面总状态开销仅约 200~KB。

	针对单星本地丢弃难以降低上游链路带宽浪费的问题，本文进一步提出多星协同在网计算与推回缓解方法 CoopSatShield。该方法通过抑制型邻居扩散协议在局部区域内聚合多星攻击证据，并结合逐跳推回机制将缓解位置从受害链路前移至上游入口，从而减少攻击流在网络中的无效占用。实验结果表明，在 $M=100$ 退化场景下，CoopSatShield 可将无效带宽占用降至无防御方案的 11.0\%，使攻击期间良性吞吐稳态维持在 0.959，控制带宽仅占 ISL 容量的 0.004\%。本文工作为大规模低轨卫星网络提供了一种兼顾低时延、低资源开销与抗退化能力的在网安全防御思路。

	% 中文关键字
	\zhkeywords{低轨卫星网络；链路洪泛攻击；在网计算；可编程数据平面}
	
	

	
	
	% 开启英文摘要
	\enabstract
	
	Large-scale Low Earth Orbit (LEO) mega-constellations, with their low latency, wide coverage, and global connectivity, are becoming an important foundation of space-ground integrated information networks. However, compared with terrestrial backbone networks, LEO constellations feature more bandwidth-limited Inter-Satellite Links (ISLs), while orbital, topological, and partial link information is highly public and predictable. This makes it easier for attackers to launch Link Flooding Attacks (LFAs) against critical links. In such attacks, a large number of distributed nodes send low-rate yet coordinated traffic to legitimate destinations, causing attack flows to converge on a target link, create congestion, and significantly degrade benign throughput and network availability.

	Existing defense mechanisms remain difficult to apply directly to LEO scenarios. On the one hand, centralized approaches relying on ground controllers or ground scrubbing centers are constrained by the latency of space-ground control loops and cannot respond promptly to millisecond-scale attack dynamics. On the other hand, programmable-data-plane solutions originally designed for terrestrial networks usually assume stable topologies and abundant hardware resources, and often rely on single features such as aggregate rate, which leads to missed detections under stealthy strategies including multi-decoy dilution, low-rate attacks, and pulsed attacks. In addition, mitigation performed only near the victim link cannot prevent wasted bandwidth consumption along upstream paths. To address these issues, this thesis investigates an in-network defense framework for resource-constrained spaceborne programmable data planes, spanning from single-node detection to multi-node cooperative pushback.

	To address the difficulty of accurately identifying stealthy LFAs from a single-satellite perspective, this thesis proposes MS-SatShield, a multi-signature in-network detection and queue-based mitigation method. MS-SatShield adopts a cascaded architecture of Top-$k$ candidate generation and multi-dimensional structural feature extraction, jointly exploiting aggregate rate, Fan-in (FI), Fan-out (FO), and Persistence (Persist) under limited state overhead to achieve robust detection against degraded attacks. Experimental results show that under a multi-decoy dilution setting ($M=100$), the F1 score of a rate-only baseline drops to 0.017, whereas incorporating FI/FO restores it to 0.908; under pulsed on-off attacks, the persistence feature further improves Recall to 0.999, while the total data-plane state overhead remains only about 200~KB.

	To address the limitation that local dropping at a single satellite cannot reduce wasted bandwidth on upstream links, this thesis further proposes CoopSatShield, a multi-satellite cooperative in-network computing and pushback mitigation method. CoopSatShield aggregates attack evidence within a local region through a suppressed neighbor-gossip protocol and combines it with a hop-by-hop pushback mechanism that moves the mitigation point from the victim link toward upstream ingress points, thereby reducing wasted bandwidth occupancy of attack traffic in the network. Experimental results show that under the degraded setting of $M=100$, CoopSatShield reduces wasted bandwidth to 11.0\% of the no-defense baseline, maintains benign throughput at a steady state of 0.959 during attacks, and consumes only 0.004\% of ISL capacity for control signaling. Overall, this work provides an in-network security defense approach for large-scale LEO satellite networks that jointly achieves low latency, low resource overhead, and robustness against attack degradation.

	% 英文关键字
	\enkeywords{LEO Satellite Networks; Link Flooding Attack; In-Network Computing; Programmable Data Plane}

	
	\tableofcontents  % 主目录，必要
	
	\listoffigures  % 图多则放，反之不放
	
	\listoftables  % 表多则放，反之不放
%	
%	\listofsymbs  % 生成主要符号表标题，需要额外维护符号表内容
		

\input{chapters/ch1_intro}

\input{chapters/ch2_background}

\input{chapters/ch3_ms_satshield}

\input{chapters/ch4_coop}

\input{chapters/ch5_conclusion}

	\acknowledgement
	\input{chapters/acknowledgement}

	
	%% 参考文献部分
	% \nocite{*}% 为了便于在示例文档中展示参考文献而设，正式撰写时需要注释掉
	% \bibliographystyle{DissertUESTC}% 自v25.03.12版本后，参考文献风格文件由用户设定的论文类型选项自动确定，无需在此手动设置
	\bibliography{ref}
	
	% 附录起始位置
%	\appendix
%	\chapter{建模证明}

	\achievement
	\input{chapters/achievement}
	
\end{document}
